---
title: "Untitled"
author: "Juan Silva"
date: "2/6/2020"
output: html_document
---

try this

what's going on?

```{r}
# For general stuff:
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)

# For ts stuff: 
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)

# For spatial stuff: 
library(sf)
library(tmap)
library(mapview)
```

# Monthly US energy consumption (renewables)

```{r}

us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

unique(us_renew$description)



```

We'll focus on consumption data. 

### Clean up data

- Convert description to all lowercase
- Only keep observations for "consumption" in "description" variable
- Remove any "total" observations from "description" variable


```{r}

renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))

```

###Convert 'yyyymm' column to a date

```{r}

renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% #coerce to tsibble `yearmonth` format
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# Want to parse the year and month? We may use this later...
renew_parsed <-renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))

```


### Look at it:

```{r}

renew_gg <- ggplot(data = renew_date, aes(x = month_sep, 
                                          y = value,
                                          group = description)) +
  geom_line(aes(color = description))

renew_gg

```

### updating my colors with paleteer

```{r}
renew_gg +
  scale_color_paletteer_d("ghibli::LaputaMedium")

```

### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

#Lets look at time series data in a couple diferent ways

```{r}
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
# renew_ts %>% gg_season(value)

ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) + 
  geom_line(aes(color = year)) +
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right")

```

# let's look at hydroelectric consumption

```{r}

hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

# Explore: 
hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
# hydro_ts %>% gg_season(value)

# OK, what if gg_season() doesn't work?
# It's just a function that uses ggplot() to do things we already know how to do in ggplot()!

ggplot(hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))

```

### Calculate summary data by time using `index_by()`

What if we want to calculate consumption by quarter? We'll use `index_by()` to tell R which "windows" to calculate a value with in. 

Quarterly:
```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~ yearquarter(.)) %>% # monthly aggregates
  summarise(
    avg_consumption = mean(value)
  )

head(hydro_quarterly)
```


### Decompose the hydro consumption ts data

First, let's check the decomposition (STL):
```{r}
# Find STL decomposition
dcmp <- hydro_ts %>%
  model(STL(value ~ season(window = 5)))

# View the components
# components(dcmp)

# Visualize the decomposed components
components(dcmp) %>% autoplot() +
  theme_minimal()

# Let's check out the residuals:
hist(components(dcmp)$remainder)

```

### Explore the ACF

```{r}

hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

```

# DANGER DANGWR DO A LOT OF READING 
MODEL FOR FORECASTING

```{r}
hydro_model <- hydro_ts %>%
  model(
    arima = ARIMA(value),
    ets = ETS(value)
  ) %>%
  fabletools::forecast(h = "2 years")

hydro_model %>% 
  autoplot(filter(hydro_ts, 
                  year(month_sep) > 2010), 
           level = NULL)


```

#lets make a world map

```{r}

# Get spatial data: 
world <- read_sf(dsn = here("data","TM_WORLD_BORDERS_SIMPL-0.3-1"), layer = "TM_WORLD_BORDERS_SIMPL-0.3") %>% clean_names()

# Quick & easy option to see those polygons (also for points, lines!)
mapview(world)

# ggplot (static)
world_base <- ggplot(data = world) +
  geom_sf(aes(fill = pop2005),
          color = NA) + 
  scale_fill_paletteer_c("viridis::viridis") +
  theme_minimal()

world_base

# Let's crop it: 
world_base +
  coord_sf(xlim = c(-20, 50), ylim = c(-40, 40), expand = FALSE)

```

